---
title: "Global GDP Analysis"
author: "Aaron Faltesek"
date: "March 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





### Order of Contents

######Introduction | 

######Folder Description | 

######Directions to run the code | 

######References  | 

#####Batch file for gather GDP data and performing analysis.




##Introduction

The folling will detail how to obtain the raw GDP and Education data sets. 

The data is from the websites https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv and 
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv

The first file is GDP data. This has 4 useful columns and several "blank" columns that iwll have to cleaned up.
The useful columns are Country Code (1), GDP Ranking (2), Full Country Name (3) and GDP Amount (4)

The second file is Education data. This data includes elaborated information about globla country data. There are over 20 columns in this data set.
The most important columns that will be used for this analysis is Long Name, Income Group, and Country Code.

Both sets of data have country code as well as superfulous records on global regions. In this analysis the two data sets will be joined/merged on the shares country code values.
Then they will have to be scrubbed to rid of unneeded regional information.

Common scrubbing necessary:
1) Remove null values
2) Remove commas from integer values for conversion and comparison
3) Remove unneeded records - Regional data
4) Remove messy headers


##Folder Structure
The folder Global_GDP_Analysis/ will contain all necessary files for this analysis
Markdown file:

-CaseStudyRMarkDownFinal_AaronFaltesek.Rmd

Main file:

Child source files:

-CleanData.R

-Question1.R

-Question2.R

-Question3.R

-Question4.R

-Question5.R

HTML Output:

-CaseStudyRMarkDownFinal_AaronFaltesek.html

#Directions for running the code.

####Libraries required
#####Install if needed below:
######install.packages("downloader")
######install.packages("ggplot2")
```{r libraries}
require(downloader)
library(downloader)

require(ggplot2)
library(ggplot2)
```

#####Set my directory where i'll be downloading files
```{r directories, include=TRUE}
setwd("D:/Users/Aaron/Documents/School/SMU/Intro_To_DataScience/Unit_8/Case_Study/Global_GDP_Analysis")
print(getwd())
```
#####Getting the raw data. This data will have to be cleaned up
```{r download, include=TRUE}
download("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv",destfile="GDP.csv")
download("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv ",destfile="Educ.csv")
```

#####This will call an R script that will clean up the data so that I can merge them together
```{r clean Data, include=TRUE}
source("CleanData.R", echo = TRUE, keep.source = TRUE)
```
#####This will contain the R code to answer question 1
```{r Question1, include=TRUE}
source("Question1.R", echo = TRUE, keep.source = TRUE)
```
#####I output my answer in the source file, but for clarify the answer to question one is that 224 rows were successfully joined.
#####----------------------------------------------------------------------

#####This will contain the R code to answer question 2
```{r Question2}
source("Question2.R", echo = TRUE, keep.source = TRUE)
```
#####----------------------------------------------------------------------

#####This will contain the R code to answer question 3
```{r Question3}
source("Question3.R", echo = TRUE, keep.source = TRUE)
```
#####----------------------------------------------------------------------

#####This will contain the R code to answer question 4
```{r Question4}
source("Question4.R", echo = TRUE, keep.source = TRUE)
```
#####----------------------------------------------------------------------

#####This will contain the R code to answer question 5
```{r Question5}
source("Question5.R", echo = TRUE, keep.source = TRUE)
```

#Conclusion
A lot can be learned from comparing GDP to income catagorization. After cleaning up the data and merging them on country code. I found that there were some missing corresponding matches. This reduced the total number of successful joins to 224. Then when I ordered the data by ascending on the GDP amount I found that St. Knitts and Nevis was the 13th value in the data set from the top. Next I found the mean rank of High income: OECD countries to be 32.97. I also found the mean rank of High income: nonOECD countries to be 91.91. The results of this question are specific to the initial rank assigned to each country. For the fourth question I found that there were significant outliers in the data that skewed the resulting plot. The US and China stand out as having exceptionally large GDPs.Lastly for question 5, I was able to compare the income groups to quantiles. I then found that of the lower middle income groups, in the top 38 counties(top quantile) there were 5 in both of these groups.

#References

https://www.r-bloggers.com/how-to-format-your-chart-and-axis-titles-in-ggplot2/

http://stackoverflow.com/questions/31149875/r-knitr-and-source-function-how-to-preserve-source-file-comments-for-html-rep

http://stackoverflow.com/questions/10966109/how-to-source-r-markdown-file-like-sourcemyfile-r

http://stackoverflow.com/questions/11728419/using-cut-and-quartile-to-generate-breaks-in-r-function